---
title: "Interactive Dashboard"
output: 
  flexdashboard::flex_dashboard:
    orientation: rows
    vertical_layout: fill
runtime: shiny
---



```{r setup, include=FALSE}
library(flexdashboard)
library(shiny)
library(ggplot2)
library(plotly)
library(dplyr)
library(readxl)
library(rlang)
library(shiny)
```

```{r}
# install.packages("readxl")
library(readxl)
mydata <- read_excel("Penguin data.xlsx")
mydata<- mydata[, -17]
miss_count <- colSums(is.na(mydata))
mydata <- na.omit(mydata)
mydata$"Culmen Length (mm)" <- as.numeric(mydata$"Culmen Length (mm)")
mydata$"Culmen Depth (mm)" <- as.numeric(mydata$"Culmen Depth (mm)")
mydata$"Flipper Length (mm)" <- as.numeric(mydata$"Flipper Length (mm)")
mydata$"Body Mass (g)" <- as.numeric(mydata$"Body Mass (g)")
mydata$"Delta 15 N (o/oo)" <- as.numeric(mydata$"Delta 15 N (o/oo)")
mydata$"Delta 13 C (o/oo)" <- as.numeric(mydata$"Delta 13 C (o/oo)")

mydata <- mydata %>% mutate(Sex = ifelse(Sex == "MALE", "MALE", "FEMALE"))


```



Page 1
======================================================================

Row {.sidebar}
-----------------------------------------------------------------------

### Chart 1: Bar chart of penguin species counts by Island:

```{r}
# create a dropdown menu for island selection
selectInput("selected_island", "Select Island:", choices = c("All", unique(mydata$Island)))

# create a reactive bar chart based on the selected island
output$bar_chart <- renderPlot({
  # filter the data by the selected island
  if (input$selected_island != "All") {
    data_filtered <- mydata %>% filter(Island == input$selected_island)
  } else {
    data_filtered <- mydata
  }
  
  # create the bar chart
  ggplot(data_filtered, aes(x = Island, fill = Species)) +
    geom_bar(position = "stack") +
    labs(title = "Penguin Species Counts by Island", x = "Island", y = "Count", fill = "Species")
})

# display the reactive bar chart
plotOutput("bar_chart")

```


### Chart 2: Scatter plot of culmen length and body mass by sex:
```{r}
# create a slider for selecting a range of body mass values
sliderInput("body_mass_range", "Select Body Mass (g) Range:",
            min = min(mydata$`Body Mass (g)`),
            max = max(mydata$`Body Mass (g)`),
            value = c(min(mydata$`Body Mass (g)`), max(mydata$`Body Mass (g)`)))

# create a reactive scatter plot based on the selected body mass range
output$scatter_plot <- renderPlot({
  # filter the data by the selected body mass range
  data_filtered <- mydata %>% filter(`Body Mass (g)` >= input$body_mass_range[1] & 
                                      `Body Mass (g)` <= input$body_mass_range[2])
  
  # create the scatter plot
  ggplot(data_filtered, aes(x = `Culmen Length (mm)`, y = `Body Mass (g)`, color = Sex)) +
    geom_point(alpha = 0.7) +
    geom_smooth(method = "lm", se = FALSE) +
    labs(title = "Relationship between Culmen Length and Body Mass by Sex", x = "Culmen Length (mm)", y = "Body Mass (g)", color = "Sex")
})

# display the reactive scatter plot
plotOutput("scatter_plot")

```


Page 2
======================================================================
Row 
-----------------------------------------------------------------------

### Chart 3: Scatter plot of culmen length and body mass by species:

```{r}
ggplot(mydata, aes(x = `Culmen Length (mm)`, y = `Body Mass (g)`, color = Species)) +
  geom_point(alpha = 0.7) +
  geom_smooth(method = "lm", se = FALSE) +
  labs(title = "Relationship between Culmen Length and Body Mass by Species", x = "Culmen Length (mm)", y = "Body Mass (g)", color = "Species")
```


### Chart 4: Plot of relationship between culmen length and culmen depth in different species.

```{r}
ggplot(mydata, aes(x = `Culmen Length (mm)`, y = `Culmen Depth (mm)`, color = Species)) +
  geom_point(alpha = 0.7) +
  geom_smooth(method = "lm", se = FALSE) +
  labs(title = "Relationship between Culmen Length and Depth by Species", x = "Culmen Length (mm)", y = "Culmen Depth (mm)", color = "Species")
```


Page 3
======================================================================

Row
-----------------------------------------------------------------------

### Chart 5: plot of changing the "Culmen Depth (mm)" by Island over time: {.storyboard}
```{r, fig.width=10, fig.height=7}
library(scales)

# convert Date Egg column to date format
mydata$`Date Egg` <- as.Date(mydata$`Date Egg`)

ggplot(mydata, aes(x = `Date Egg`, y = `Culmen Depth (mm)`)) +
  geom_point(alpha = 0.2) +
  geom_smooth() +
  labs(title = "Trend of Culmen Depth (mm) by Island over Time", x = "Date", y = "Culmen Depth (mm)") +
  facet_grid(. ~ Island) +
  scale_x_date(date_breaks = "1 year", date_labels = "%Y")

```


Page 4
======================================================================

Row
-----------------------------------------------------------------------

### Chart 6: Visualization for exploring the relationship between body mass and flipper length.
```{r}
# create scatter plot of body mass vs. flipper length
ggplot(mydata, aes(x = `Flipper Length (mm)`, y = `Body Mass (g)`, color = Species)) +
  geom_point(size = 2) +
  labs(x = "Flipper Length (mm)", y = "Body Mass (g)", color = "Species") +
  ggtitle("Body Mass vs. Flipper Length by Species")
```


### Chart 7: Visualization for exploring the trend of culmen length by penguin species and sex over time.
```{r}
library(scales)
library(stringr)

# convert Date Egg column to date format
mydata$`Date Egg` <- as.Date(mydata$`Date Egg`)

# extract first word of species names
mydata$short_species <- str_extract(mydata$Species, "^\\w+")

# create scatter plot of culmen length vs. year, faceted by species and sex
ggplot(mydata, aes(x = `Date Egg`, y = `Culmen Length (mm)`, color = Sex)) +
  geom_point(alpha = 0.5) +
  geom_smooth(se = FALSE) +
  facet_grid(short_species ~ ., labeller = label_parsed) +
  labs(x = "Year", y = "Culmen Length (mm)", color = "Sex") +
  ggtitle("Trend of Culmen Length by Species and Sex over Time") +
  scale_x_date(date_breaks = "3 months", date_labels = "%b %Y")
```


Page 5
======================================================================


Row
-----------------------------------------------------------------------

### Chart 8: Relationship between Body Mass and Delta 15 N
```{r}
ggplot(mydata, aes(x = `Body Mass (g)`, y = `Delta 15 N (o/oo)`)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  labs(x = "Body Mass (g)", y = "Delta 15 N (o/oo)", 
       title = "Relationship between Body Mass and Delta 15 N") +
  theme_bw()

```



### Chart 9: Relationship between Body Mass and Delta 13 C (o/oo)
```{r}
library(ggplot2)

ggplot(mydata, aes(x = `Body Mass (g)`, y = `Delta 13 C (o/oo)`)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  labs(x = "Body Mass (g)", y = "Delta 13 C (o/oo)", 
       title = "Relationship between Body Mass and Delta 13 C (o/oo)") +
  theme_bw()
```

